Plots

### 2025-05-23 manuscript/manuscript.tex
Added data profile by cod

>>>>

\begin{sidewaystable}
\caption{Case study data by cause of death.}\label{tab-data-cod}
\begin{tabular*}{\textheight}{@{\extracolsep\fill}lllll}
    \toprule%
    Age Group  &  Cause of Death (COD)  &  Female  &  Male  &  Total  \\
    \midrule
    \makecell[l]{\\Adult (n=3826, 55.1\%, 18 CODs)\\Adult Female (n=1681, 43.9\%)\\Adult Male (n=2145, 56.1\%)}  &  Acute Respiratory Infections  &  48 (45.7\%)  &  57 (54.3\%)  &  105 (2.7\%)  \\ 
        &  Cancers  &  32 (65.3\%)  &  17 (34.7\%)  &  49 (1.3\%)  \\ 
        &  Chronic Respiratory Diseases  &  29 (38.7\%)  &  46 (61.3\%)  &  75 (2\%)  \\ 
        &  Diabetes Mellitus  &  14 (51.9\%)  &  13 (48.1\%)  &  27 (0.7\%)  \\ 
        &  Diarrhoeal Diseases  &  102 (49.8\%)  &  103 (50.2\%)  &  205 (5.4\%)  \\ 
        &  Ill-Defined  &  56 (47.9\%)  &  61 (52.1\%)  &  117 (3.1\%)  \\ 
        &  Ischemic Heart Disease  &  89 (53\%)  &  79 (47\%)  &  168 (4.4\%)  \\ 
        &  Liver And Alcohol Related Diseases  &  58 (45.3\%)  &  70 (54.7\%)  &  128 (3.3\%)  \\ 
        &  Malaria  &  372 (46.6\%)  &  427 (53.4\%)  &  799 (20.9\%)  \\ 
        &  Maternal Conditions  &  130 (100\%)  &  N/A  &  130 (3.4\%)  \\ 
        &  Other Cardiovascular Diseases  &  59 (55.1\%)  &  48 (44.9\%)  &  107 (2.8\%)  \\ 
        &  Other Noncommunicable Diseases  &  160 (38.6\%)  &  254 (61.4\%)  &  414 (10.8\%)  \\ 
        &  Other Injuries  &  83 (23.2\%)  &  274 (76.8\%)  &  357 (9.3\%)  \\ 
        &  Road And Transport Injuries  &  73 (29.1\%)  &  178 (70.9\%)  &  251 (6.6\%)  \\ 
        &  Stroke  &  147 (44.4\%)  &  184 (55.6\%)  &  331 (8.7\%)  \\ 
        &  Suicide  &  N/A  &  3 (100\%)  &  3 (0.1\%)  \\ 
        &  Tuberculosis  &  54 (31.6\%)  &  117 (68.4\%)  &  171 (4.5\%)  \\ 
        &  UnspecifiedInfections  &  175 (45\%)  &  214 (55\%)  &  389 (10.2\%)  \\ 
    \makecell[l]{\\Child (n=2636, 38\%, 9 CODs)\\Child Female (n=1290, 48.9\%)\\Child Male (n=1346, 51.1\%)}  &  CongenitalAnomalies  &  1 (100\%)  &  N/A  &  1 (0\%)  \\ 
        &  Diarrhoeal Diseases  &  79 (45.1\%)  &  96 (54.9\%)  &  175 (6.6\%)  \\ 
        &  Epilepsy, Leukaemia, And Other Noncommunicable Diseases  &  61 (53.5\%)  &  53 (46.5\%)  &  114 (4.3\%)  \\ 
        &  Ill-Defined  &  34 (48.6\%)  &  36 (51.4\%)  &  70 (2.7\%)  \\ 
        &  Injuries  &  51 (37.8\%)  &  84 (62.2\%)  &  135 (5.1\%)  \\ 
        &  Malaria  &  680 (49.2\%)  &  702 (50.8\%)  &  1382 (52.4\%)  \\ 
        &  Nutritional Deficiencies  &  7 (63.6\%)  &  4 (36.4\%)  &  11 (0.4\%)  \\ 
        &  Other Infections  &  338 (50.7\%)  &  329 (49.3\%)  &  667 (25.3\%)  \\ 
        &  Pneumonia  &  39 (48.1\%)  &  42 (51.9\%)  &  81 (3.1\%)  \\ 
    \makecell[l]{\\Neonate (n=477, 6.9\%, 7 CODs)\\Neonate Female (n=227, 47.6\%)\\Neonate Male (n=250, 52.4\%)}  &  Birth Asphyxia And Birth Trauma  &  38 (36.9\%)  &  65 (63.1\%)  &  103 (21.6\%)  \\ 
        &  Congenital Anomalies  &  2 (100\%)  &  N/A  &  2 (0.4\%)  \\ 
        &  Ill-Defined  &  11 (47.8\%)  &  12 (52.2\%)  &  23 (4.8\%)  \\ 
        &  Neonatal Infections  &  49 (49.5\%)  &  50 (50.5\%)  &  99 (20.8\%)  \\ 
        &  Other  &  2 (40\%)  &  3 (60\%)  &  5 (1\%)  \\ 
        &  Prematurity And Low Birthweight  &  39 (53.4\%)  &  34 (46.6\%)  &  73 (15.3\%)  \\ 
        &  Stillbirth  &  86 (50\%)  &  86 (50\%)  &  172 (36.1\%)  \\
        \botrule
\end{tabular*}
\end{sidewaystable}

\begin{sidewaystable}
    \caption{Case study data by age range.}\label{tab-data-age}
    \begin{tabular*}{\textheight}{@{\extracolsep\fill}lllll}
        \toprule%
        Age Group  &  Cause of Death (COD)  &  Female  &  Male  &  Total  \\
        \midrule
        \makecell[l]{\\Adult (n=3826, 55.1\%, 18 CODs)\\Adult Female (n=1681, 43.9\%)\\Adult Male (n=2145, 56.1\%)}  &  Acute Respiratory Infections  &  48 (45.7\%)  &  57 (54.3\%)  &  105 (2.7\%)  \\ 
            &  Cancers  &  32 (65.3\%)  &  17 (34.7\%)  &  49 (1.3\%)  \\ 
            &  Chronic Respiratory Diseases  &  29 (38.7\%)  &  46 (61.3\%)  &  75 (2\%)  \\ 
            &  Diabetes Mellitus  &  14 (51.9\%)  &  13 (48.1\%)  &  27 (0.7\%)  \\ 
            &  Diarrhoeal Diseases  &  102 (49.8\%)  &  103 (50.2\%)  &  205 (5.4\%)  \\ 
            &  Ill-Defined  &  56 (47.9\%)  &  61 (52.1\%)  &  117 (3.1\%)  \\ 
            &  Ischemic Heart Disease  &  89 (53\%)  &  79 (47\%)  &  168 (4.4\%)  \\ 
            &  Liver And Alcohol Related Diseases  &  58 (45.3\%)  &  70 (54.7\%)  &  128 (3.3\%)  \\ 
            &  Malaria  &  372 (46.6\%)  &  427 (53.4\%)  &  799 (20.9\%)  \\ 
            &  Maternal Conditions  &  130 (100\%)  &  N/A  &  130 (3.4\%)  \\ 
            &  Other Cardiovascular Diseases  &  59 (55.1\%)  &  48 (44.9\%)  &  107 (2.8\%)  \\ 
            &  Other Noncommunicable Diseases  &  160 (38.6\%)  &  254 (61.4\%)  &  414 (10.8\%)  \\ 
            &  Other Injuries  &  83 (23.2\%)  &  274 (76.8\%)  &  357 (9.3\%)  \\ 
            &  Road And Transport Injuries  &  73 (29.1\%)  &  178 (70.9\%)  &  251 (6.6\%)  \\ 
            &  Stroke  &  147 (44.4\%)  &  184 (55.6\%)  &  331 (8.7\%)  \\ 
            &  Suicide  &  N/A  &  3 (100\%)  &  3 (0.1\%)  \\ 
            &  Tuberculosis  &  54 (31.6\%)  &  117 (68.4\%)  &  171 (4.5\%)  \\ 
            &  UnspecifiedInfections  &  175 (45\%)  &  214 (55\%)  &  389 (10.2\%)  \\ 
        \makecell[l]{\\Child (n=2636, 38\%, 9 CODs)\\Child Female (n=1290, 48.9\%)\\Child Male (n=1346, 51.1\%)}  &  CongenitalAnomalies  &  1 (100\%)  &  N/A  &  1 (0\%)  \\ 
            &  Diarrhoeal Diseases  &  79 (45.1\%)  &  96 (54.9\%)  &  175 (6.6\%)  \\ 
            &  Epilepsy, Leukaemia, And Other Noncommunicable Diseases  &  61 (53.5\%)  &  53 (46.5\%)  &  114 (4.3\%)  \\ 
            &  Ill-Defined  &  34 (48.6\%)  &  36 (51.4\%)  &  70 (2.7\%)  \\ 
            &  Injuries  &  51 (37.8\%)  &  84 (62.2\%)  &  135 (5.1\%)  \\ 
            &  Malaria  &  680 (49.2\%)  &  702 (50.8\%)  &  1382 (52.4\%)  \\ 
            &  Nutritional Deficiencies  &  7 (63.6\%)  &  4 (36.4\%)  &  11 (0.4\%)  \\ 
            &  Other Infections  &  338 (50.7\%)  &  329 (49.3\%)  &  667 (25.3\%)  \\ 
            &  Pneumonia  &  39 (48.1\%)  &  42 (51.9\%)  &  81 (3.1\%)  \\ 
        \makecell[l]{\\Neonate (n=477, 6.9\%, 7 CODs)\\Neonate Female (n=227, 47.6\%)\\Neonate Male (n=250, 52.4\%)}  &  Birth Asphyxia And Birth Trauma  &  38 (36.9\%)  &  65 (63.1\%)  &  103 (21.6\%)  \\ 
            &  Congenital Anomalies  &  2 (100\%)  &  N/A  &  2 (0.4\%)  \\ 
            &  Ill-Defined  &  11 (47.8\%)  &  12 (52.2\%)  &  23 (4.8\%)  \\ 
            &  Neonatal Infections  &  49 (49.5\%)  &  50 (50.5\%)  &  99 (20.8\%)  \\ 
            &  Other  &  2 (40\%)  &  3 (60\%)  &  5 (1\%)  \\ 
            &  Prematurity And Low Birthweight  &  39 (53.4\%)  &  34 (46.6\%)  &  73 (15.3\%)  \\ 
            &  Stillbirth  &  86 (50\%)  &  86 (50\%)  &  172 (36.1\%)  \\
            \botrule
    \end{tabular*}
    \end{sidewaystable}

### 2025-05-23 src/plots.Rmd
Added data profile by cod

```{r, echo=FALSE, eval=FALSE}
#install.packages("tidyverse")
#install.packages("dplyr")
#install.packages("ggrepel")
#install.packages("patchwork")
```

@@

```{r}
library(tidyverse)
library(dplyr)
library(ggrepel)
library(patchwork)
```

@@

>>>>

```{r, echo=FALSE, eval=FALSE}
#install.packages("openVA")
#install.packages("tidyverse")
#install.packages("dplyr")
#install.packages("ggrepel")
#install.packages("patchwork")
#install.packages("glue")
```

@@

```{r}
library(tidyverse)
library(dplyr)
library(ggrepel)
library(patchwork)
library(glue)
```

@@

# Tables

# Get n records by age group
tab_agegroup <- df %>%
    group_by(age) %>%
    summarize(
        n = n(),
        n_cod_uniq = n_distinct(physician_cghr10)
    )

# Get n records by age group and sex
tab_agegroup_sex <- df %>%
    group_by(age, sex) %>%
    summarize(n_sex = n()) %>%
    pivot_wider(
        names_from = sex,
        values_from = n_sex,
        names_glue = "n_{str_to_lower(sex)}"
    )

## Records by COD

# Get n records by cod
tab_cod_n <- df %>%
    group_by(age, physician_cghr10) %>%
    summarize(n_cod = n())

# Get n records by cod and sex
tab_cod_sex <- df %>%
    group_by(age, physician_cghr10, sex) %>%
    summarize(n_cod_sex = n()) %>%
    pivot_wider(
        names_from = sex,
        values_from = n_cod_sex,
        names_glue = "n_cod_{str_to_lower(sex)}"
    )

# Combine all
tab_cod <- tab_cod_n %>%
    left_join(tab_agegroup, by = "age") %>%
    left_join(tab_agegroup_sex, by = "age") %>%
    left_join(tab_cod_sex, by = c("age", "physician_cghr10"))

# Calc perc
tab_cod <- tab_cod %>%
    mutate(
        n_perc = round(n / nrow(df) * 100, 1),
        n_male_perc = round(n_male / n * 100, 1),
        n_female_perc = round(n_female / n * 100, 1),
        n_cod_perc = round(n_cod / n * 100, 1),
        n_cod_male_perc = round(n_cod_male / n_cod * 100, 1),
        n_cod_female_perc = round(n_cod_female / n_cod * 100, 1)
    ) %>%
    ungroup

# Add labels to combined
tab_cod <- tab_cod %>%
    mutate(
        age = if_else(age == "neo", "Neonate", str_to_title(age)),
        `Age Group` = glue(
            "{age} (n={n}, {n_perc}%, {n_cod_uniq} CODs)\n{age} Female (n={n_female}, {n_female_perc}%)\n{age} Male (n={n_male}, {n_male_perc}%)"
        ),
        `Cause of Death (COD)` = glue(
            "{str_to_title(physician_cghr10)}"
        ),
        `Female` = glue("{n_cod_female} ({n_cod_female_perc}%)"),
        `Male` = glue("{n_cod_male} ({n_cod_male_perc}%)"),
        `Total` = glue("{n_cod} ({n_cod_perc}%)")
    ) %>%
    mutate(across(
        c(`Female`, `Male`, `Total`),
        ~ if_else(.x == "NA (NA%)", NA, .x)
    ))

# Save table
write_csv(tab_cod, "../data/healsl_rd1to2_cod_profile_v1.csv", na = "")

# Display table
tab_cod
```

# Prepare and format the data
tab_cod_format <- tab_cod %>%
    ungroup %>%
    select(`Age Group`, `Cause of Death (COD)`, Female, Male, Total) %>%
    mutate(
        `Age Group` = str_replace_all(`Age Group`, "\\\\n|\\n", "<br>"),
        `Cause of Death (COD)` = str_replace_all(`Cause of Death (COD)`, "\\\\n|\\n", "<br>")
    ) %>%
    group_by(`Age Group`) %>%
    mutate(`Age Group` = ifelse(row_number() == 1, `Age Group`, "")) %>%
    ungroup()

# Build the table
tab_cod_html <- tab_cod_format %>%
    kable(format = "html", escape = FALSE, align = "l") %>%
    kable_styling(full_width = FALSE, position = "left", html_font = "Arial") %>%
    row_spec(0, extra_css = "border-top: 2px solid; border-bottom: 1px solid;") %>%     # thick top border
    row_spec(nrow(df_selected), extra_css = "border-bottom: 2px solid;")
tab_cod_html
```

## Records by Age

# Get n records by age
tab_age_n <- df %>%
    group_by(age, physician_cghr10) %>%
    summarize(n_cod = n())
